---
title: "Week 6 Assignment-Project 2"
output: html_document
author: "Fl0wControl"
date: "OCtober 9, 2016"
---
The Data Set: Demographic_Statistics_By_Zip_Code.csv
Suggested by: Dhananjay Kumar
Analysis:  "One can use this dataset for targeted Marketing Campaign based on Demography. For example, a manufacturer of Sporting Goods or a Sport Store can choose a county for their campaign based on residents age. Another example would be to do targeted campaign on the basis of second/native language."

Note: Since no age was given, nor was second or native language, I decided to see what the demographics of the data set looked like in terms of gender, ethnicity, and citizenship status.


Load the required R packages:
```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
```

Load in the data from a hosted source, and create a function to gather the data:
```{r}
nyDemographics <- read.table("https://raw.githubusercontent.com/Fl0wControl/Data607/master/Week6_Assignment/Demographic_Statistics_By_Zip_Code.csv", header = TRUE, sep=",")

createGatheredData <- function(dataSet, keyCol, valueCol, gatherCol){
  gathered <- gather_(dataSet, keyCol, valueCol, gatherCol)
  return (gathered)
}
```

I am curious to see what the makeup of the data is as far as the total population goes. I want to see how many total males and females represent the data, so we shall create a
small representation of male and female respondants relative to the total surveyed.
```{r}
maleVsFemale <- nyDemographics %>% summarise(TotalMales=sum(COUNT.MALE), TotalFemales=sum(COUNT.FEMALE), TotalSurveyed=sum(COUNT.PARTICIPANTS), PercentMale=(TotalMales/TotalSurveyed), PercentFemale=(TotalFemales/TotalSurveyed))

maleVsFemale
```
Males make up 1738 of those surveyed, while females makeup up 2430 of the total 4168 surveyed.

Now we want to see the total ethnica makeup of the data. Is one group truly a minority?
```{r}
ethnicDiversity <- nyDemographics %>% summarise(
TotalPacificIslander=sum(COUNT.PACIFIC.ISLANDER),
TotalHispanic=sum(COUNT.HISPANIC.LATINO),
TotalAmericanIndian=sum(COUNT.AMERICAN.INDIAN),
TotalAsian=sum(COUNT.ASIAN.NON.HISPANIC),
TotalWhite=sum(COUNT.WHITE.NON.HISPANIC),
TotalBlack=sum(COUNT.BLACK.NON.HISPANIC),
TotalOther=sum(COUNT.OTHER.ETHNICITY))

ethnicDiversity
```
Based on the data, we can conclude that whites do make up the majority of the population, while American Indian's are the minority with the fewest surveyed.

To see the citizen status of those surveyed in the demographics data, we need to summarise those who participated. After that, we can get a clearer picture of who
is legally in the country or not.
```{r}
citizenStatus <- nyDemographics %>% summarise(TotalCitizenStatus = sum(COUNT.US.CITIZEN),
TotalOtherStatus = sum(COUNT.OTHER.CITIZEN.STATUS),
TotalUnknownStatus = sum(COUNT.CITIZEN.STATUS.UNKNOWN))

citizenStatus
```
Surprisingly, 4052 responded with citizenship status, while only 11 responded with a status of "other"