---
title: "Week7 Assignment"
output: html_document
---

Load the required libraries into R. XML is for parsing XML data, jsonlite will handle any JSON inputs, RCurl is for retrieving data from any URL, and plyr will be used in manipulating and cleaning up data.
```{r}
library(XML)
library(jsonlite)
library(RCurl)
library(plyr)
```

Take the data from the XML format, and get it into a data frame. Ldply returns a data frame, and xmlToList is good for handling responses from a REST call. 
```{r}
bookXml <-ldply(xmlToList(getURL("https://raw.githubusercontent.com/Fl0wControl/Data607/master/Week7_Assignment/books.xml")), data.frame)

str(bookXml)
```

To get the HTML table read into a data frame, we need to get the data from the URL with getURL, as there are some known issues with the readHtmlTable method. 
```{r}
htmlData <- getURL("https://raw.githubusercontent.com/Fl0wControl/Data607/master/Week7_Assignment/books.html")
bookHtml <- readHTMLTable(htmlData, which=1)

str(bookHtml)
```

The jsonlite package has a simple one liner that runs out, gets the data, and turns it into a data frame.
```{r}
bookJson <- fromJSON("https://raw.githubusercontent.com/Fl0wControl/Data607/master/Week7_Assignment/books.json")

str(bookJson)
```

The data frames are not the same. The html and json data frames have 7 variables, while the xml data frame has 11. This is because of the multiple authors for "Learning Spark". However, all have 3 observations.