---
title: "Week 6 Assignment-Project 2"
output: html_document
author: "Fl0wControl"
date: "OCtober 9, 2016"
---

The Data Set: 201606-citibike-tripdata.csv (a small sample of it)
Suggested by: Sharon Morris
Analysis:  The data can be analyzed to create a profile of they typical Citibike users/subscriber. The year can be converted to an age and converted to a categorical variable. Start and stop times can be used to calculate the average time each bike is used.

Load the required R packages, read in the data, and analyze the structure of the data:
```{r}
citiBike <- read.table("https://raw.githubusercontent.com/Fl0wControl/Data607/master/Week6_Assignment/short_citi_bike.csv", header = TRUE, sep=",")

str(citiBike)
```
In order to analyze the age of the sample, we need to first get the information.

```{r}
citiAges <- citiBike$birth.year

citiAges
```

We see there are NA values, so we need to remove them from the data when calculating an age.
```{r}
ages <- na.exclude(2016-citiAges)
summary(ages)
```
From this, we can see the mean age is 35.42, and the median is 33. Interestingly, the youngest to ride is 19 and oldest is 71 years of age.

The next requested analysis was to find the start and stop times, and find the average time each bike was used. Being there is a "tripduration" variable, we can just convert this into minutes, as the total is in seconds.  I found this by taking the first stop time, and last stop time, and finding the difference. As they are already in minute format, I can find the difference.

```{r}
head(citiBike$starttime)
head(citiBike$stoptime)
```
The first data point is "6/1/2016 00:24:48" in stop time, and we can subtract the start time point of "6/1/2016 00:00:18" by converting the items into seconds.
```{r}
(((24*60)+48)-(18))
```
This is equal to the first item in "tripduration", which is 1470.
```{r}
head(citiBike$tripduration)
```

To find the average use time, we can perform summary command against the trip duration, divided by 60. This gives us the min time of 1.48 minutes, a mean time of 11.120, and a max time of 257 minutes.
```{r}
summary(citiBike$tripduration/60)
```
