---
title: "Week 3 Assignment"
date: "September 16, 2016"
output: html_document
---

```{r}
library(stringr)
```

## 3. Copy the introductory example. The vector name stores the extracted names.
a. Use the tools of this chapter to rearrange the vector so that all elements conform to the standard 

```{r}
originalData <-"555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"
name <- unlist(str_extract_all(originalData, "[[:alpha:]., ]{2,}"))
removeAbbrv <- str_replace_all(name, "C. ","")
removeAbbrv <- str_replace_all(removeAbbrv, c("Rev. ", "Dr. "),"")
fixburns <- sub('^(.*), montgomery', 'montgomery \\1', tolower(removeAbbrv))
fixhomer <- sub('^(.*), homer', 'homer \\1', tolower(fixburns))
camelCase <- str_replace_all(fixhomer, " ","_")
```

b. Construct a logical vector indicating whether a character has a title (i.e., Rev. and Dr.).
```{r}
str_detect(name,c("Rev.","Dr."))
str_extract(name,c("Rev.","Dr."))
```

c . Construct a logical vector indicating whether a character has a second name.
```{r}
str_detect(fixhomer, "(.+?) (.+?)")
str_extract(fixhomer, "(.+?) (.+?)")
```

## 4.Describe the types of strings that conform to the following regular expressions and construct an example that is matched by the regular expression.

a. [0-9]+\\$  - this matches any number plus the dollar sign (which is escaped)
```{r}
item <- c("989898$", "something", "2323")
str_detect(item, "[0-9]+\\$")
str_extract(item, "[0-9]+\\$")
```

b. \\b[a-z]{1,4}\\b - matches up to four letters
```{r}
item <- c("abcd","a great file.txt format", "something", "2323")
str_detect(item, "\\b[a-z]{1,4}\\b")
str_extract(item, "\\b[a-z]{1,4}\\b")
```

c. \\d{2}/\\d{2}/\\d{4} - matches digits like a date format xx/xx/xxxx
```{r}
item <- c("a great file.txt format", "111/111/11111", "01/01/111", "01/01/1234")
str_detect(item, "\\d{2}/\\d{2}/\\d{4}")
str_extract(item, "\\d{2}/\\d{2}/\\d{4}")
```

d. <(.+?)>.+?</\\1>- matches anything looking like an html tag, anything in the middle, and anything with html tag format matching the first group
```{r}
item <- c("<body> I have a great body </body>", " <no > you dont < no>", "1212")
str_detect(item, "<(.+?)>.+?</\\1>")
str_extract(item, "<(.+?)>.+?</\\1>")
```
